---
title: "CUNY DT 624"
author: 'Group 1: Avraham Adler, Vishal Arora, Samuel Bellows, Austin Chan'
date: "Summer 2020"
output:
  word_document:
    toc: yes
    toc_depth: 4
subtitle: Project 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
library(openxlsx)
library(caret)
library(data.table)
```

# Agenda
This report will discuss our team's building a model to predict the PH of a
formulation given the presence and ratio of key ingredients. We understand that
there will be a varied audience for this report. Therefore, the first section of
the report will contain a high-level overview of the modeling process and
decisions together with the findings. There will be a more technical appendix
with a deeper dive into the process. Computer code will be made available upon
request.

```{r dataIntake, include=FALSE}
trainD <- read.xlsx('./project2data/StudentData - TO MODEL.xlsx',
                    colNames = TRUE, sep.names = ' ')
testD <- read.xlsx('./project2data/StudentEvaluation- TO PREDICT.xlsx',
                    colNames = TRUE, sep.names = ' ')
setDT(trainD)
setDT(testD)
trainX <- trainD[, !"PH"]
trainY <- trainD[, PH]
numObs <- length(trainY)
numVar <- dim(trainX)[[2]]
```

# Research Problem
The company has provided data on `r numObs` formulations of `r numVar`
components and their resulting PH values. From this data, the research team was
asked to build a model allowing PH prediction based on the formulation *prior*
to actually implementing the recipe.

## Vocabulary
In the ensuing report, we will refer to the formulations *with* their applicable
PH values as the **training** set, for it is on this data we intend to *train*
the model.

The term **features** is used to describe the elements of the formulation which
will be used in prediction. In statistical terms these are also known as the
*independent variables*.

The term **target** is used to describe the element for which the model is being
built---in this case, the `PH`. In statistical terms this is also known as the
*dependent variables*.

## Data Weaknesses
Unfortunately, there were some issues with measurements. There are 




# Technical Appendix
## Missing Data





