---
title: "CUNY DT 624"
subtitle: "Project 1"
author: "Group 1: Avraham Adler, Vishal Arora, Samuel Bellows, Austin Chan"
date: "Summer 2020"
output:
  word_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(data.table)
```

# Introduction
The purpose of this report is to describe the analysis performed in order to
forecast the following series, each for 140 periods.

 * S01: Forecast Var01 and Var02
 * S02: Forecast Var02 and Var03
 * S03: Forecast Var05 and Var07
 * S04: Forecast Var01 and Var02
 * S05: Forecast Var02 and Var03
 * S06: Forecast Var05 and Var07
 
The forecasts will be provided in a seperate Excel document. The code for the
analysis can be found in the attached R markdown document.

# Executive Summary

# Analysis
## Data Preprocessing
The data was converted to CSV, read into R, and then split into seperate files
by group.
```{r datapreprocess, include=FALSE}
# Read in Data
DT <- read_xls('./project1data/Data Set for class.xls')
# Set as data table
setDT(DT)
# Seperate out groups
n <- length(unique(DT$group))
for (i in seq_len(n)) {
  gName <- paste0('S0', i)
  assign(gName, DT[group == gName])
}
```

## Data Investigation
Each of the seperate series was investigated for correlation, autocorrelation,
and stationarity.
